A-maze of Natural Stories: Texts are comprehensible using the Maze task

### 2022-10-31 Papers/Paper/manuscript.Rmd
Roger in the middle of changes

The Maze task has much less face-validity than eye-tracking or even SPR; repeated forced-choice selections does not look very similar to normal reading. Despite this, @forsterMazeTaskMeasuring2009 report that "At a phenomenological level, participants typically report that they feel as if they are reading the sentence relatively
naturally and that the correct alternative seems to “leap out” at them, so that they do not have to inspect the incorrect
alternative very carefully, if at all." This suggests that the Maze task is not about conscious comparison between two potential continuations and instead relies on the same language processing facilities shared by other methods. Maze is not trying to study the process of normal reading, but instead to get underlying language processing, using an artificial method to slow down and localize language processing.

The downside of G-maze is that materials are effort-intensive to construct because of the need to select infelicitous words as distractors for each spot of each sentence. This burdensome preparation may explain why the Maze task was not widely adopted. @boyceMazeMadeEasy2020 demonstrated a way to automatically generate Maze distractors by using language models from Natural Language Processing to find words that are high surprisal in the context of the target sentence, and thus likely to be judged infelicitous by human readers. @boyceMazeMadeEasy2020 call Maze with automatically generated distractors A-maze. In a comparison, A-maze distractors had similar results to the hand-generated G-maze distractors from @witzelComparisonsOnlineReading2012a and A-maze outperformed L-maze and an SPR control in detecting and localizing expected slowdown effects. @sloggettAmazeAnyOther2020 also found that A-maze and G-maze distractors yielded similar results on a disambiguation paradigm.

@@

Frequency is often found to be a predictor of reading time in SPR and eye-tracking experiments. It is unclear why we find frequency effects on RT when surprisal effects are accounted for. In their discussion, @smith-levy:2008 posit that the role of frequency could diminish or disappear as measurements of surprisal improve (as they have over the past decade). Additionally, @smith-levy:2008 hypothesize that the effects of frequency (in addition to those of surprisal) may be the result of quick, heuristic responses to the words based on the more easily available unigram frequency before more fine-grained context-sensitive surprisal is available. An alternative is that frequency effects don't actually exist, but are an artifact of how RT data is analysed [@shainLargescaleStudyEffects2019].

>>>>

The Maze task has less face-validity than eye-tracking or even SPR; repeated forced-choice selections does not seem very similar to normal reading. Despite this, @forsterMazeTaskMeasuring2009 report that "At a phenomenological level, participants typically report that they feel as if they are reading the sentence relatively
naturally and that the correct alternative seems to “leap out” at them, so that they do not have to inspect the incorrect
alternative very carefully, if at all." This suggests that the Maze task may rely on the same language processing facilities tapped into by other reading methods. Thus, using Maze may not be the best paradigm for studying the process of normal reading, but may be perfectly good or even superior for getting at underlying language processing.

However, G-maze materials are effort-intensive to construct because of the need to select infelicitous words as distractors for each spot of each sentence. This burdensome preparation may explain why the Maze task was not widely adopted. @boyceMazeMadeEasy2020 demonstrated a way to automatically generate Maze distractors by using language models from Natural Language Processing to find words that are high surprisal in the context of the target sentence, and thus likely to be judged infelicitous by human readers. @boyceMazeMadeEasy2020 call Maze with automatically generated distractors A-maze. In a comparison, A-maze distractors had similar results to the hand-generated G-maze distractors from @witzelComparisonsOnlineReading2012a and A-maze outperformed L-maze and an SPR control in detecting and localizing expected slowdown effects. @sloggettAmazeAnyOther2020 also found that A-maze and G-maze distractors yielded similar results on a disambiguation paradigm.

@@

The role of word frequency is worth special note. Although the facilitative effect of word frequency on reading measures has been known for decades, it is theoretically unclear why this effect should exist in any model that is sensitive to conditional word probability:

@smith-levy:2008 noted this puzzle and posited that the role of frequency could diminish or disappear as measurements of surprisal improve (as they have over the past decade), or alternatively that the effects of frequency (in addition to those of surprisal) may be the result of quick, heuristic responses to the words based on the more easily available unigram frequency before more fine-grained context-sensitive surprisal is available. An alternative is that frequency effects don't actually exist, but are an artifact of how RT data is analysed [@shainLargescaleStudyEffects2019].
